---
title: "p8105_hw3_yy3563"
author: "Yifei Yu"
date: "2024-10-12"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
library(haven)
```


## Problem 1

```{r}
library(p8105.datasets)
data("ny_noaa")
```

```{r}
head(ny_noaa)
dim(ny_noaa)
str(ny_noaa)
summary(ny_noaa)
```

The `ny_noaa` dataset includes variables such as `id`, `date`, `prcp`, `snow`, `snwd`, `tmax`, and `tmin`. There are 2595176 observations in the dataset. There are `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. `id`, `tmax`, and `tmin` are characters. `prcp`, `snow`, and `snwd` are integers.

```{r}
sapply(ny_noaa, function(x) sum(is.na(x)))
sum(is.na(ny_noaa))
```

We can see that `id` and `date` have no missing value, while `prcp` has 145838 missing value, `snow` has 381221, `snwd` has 591786, `tmax` has 1134358, and `tmin` has 1134420 missing value. The overall number of missing value is 3387623.
Both `tmax` and `tmin` have a lot of missing value, which may lead to issues when analyzing data.







